PyQBench: a Python library for benchmarking gate-based quantum computers

### 2023-03-21 txt/manuscript.tex
Address the remaining todos

\todo[inline]{Konrad + ta nazwa sekcji}
\subsubsection{Importance of the optimal discrimination scheme}
In principle, the schemes described in the previous section could be used with any choice of
$\ket{\psi_0}$ and final measurements $\PP_{V_i}$. However, we argue that it is best to choose those
components in such a way that they maximize the probability of correct discrimination. To see that,
suppose that some choice of $\ket{\psi_0}, \PP_{V_0}, \PP_{V_1}$ yields the theoretical upper bound
of discriminating between two measurements of one, i.e. on a perfect quantum computer you will always
make a correct guess. Then, on real hardware, we might obtain any empirical value in range $\left[\frac{1}{2},
1\right]$. On the other hand, if we choose the components of our scheme such that the successful
discrimination probability is only $\frac{3}{5}$, the possible range of empirically obtainable probabilities
is only $\left[\frac{1}{2}, \frac{3}{5}\right]$. Hence, in the second case, the discrepancy between theoretical and empirical
results will be less pronounced.

\subsubsection{Constructing optimal discrimination scheme}
The celebrated result by Helstrom~\cite{helstrom1976quantum} gives the optimal  probability of correct discrimination between two quantum measurements, $\PP$  and $\mathcal{Q}$,
in terms of their distance with the use of the diamond norm
\begin{equation}
p_{\text{succ}}(\PP, \mathcal{Q}) =  \frac12 + \frac14 \| \PP - \mathcal{Q} \|_\diamond,
\end{equation}
where
\begin{equation}
\|\PP - \QQ\|_\diamond = \max_{\| \ket{\psi}\|_1=1} \| \left( (\PP - \QQ) \otimes \1\right) (\proj{\psi}) \|_1.
\end{equation}
The quantum state $\ket{\psi_0}$ which maximizes the diamond norm is called the
\emph{discriminator}. Furthermore,  thanks to the proof of the Holevo-Helstrom theorem, it is
possible to construct corresponding unitaries $V_0$, $V_1$ to create the optimal
discrimination strategy. For brevity, we do not describe this procedure here. Instead, we refer the
interested reader to \cite{watrous}.

@@

\subsection{Using user-defined measurement with \texttt{qbench} package}
In this example, we will demonstrate how \texttt{qbench} package can be used with user--defined
measurement. For this purpose, we will use $U = H$ (the Hadamard gate). The detailed calculations
that lead to the particular form of the discriminator and final measurements can be found in
\ref{app:hadamard}.
\noindent
The explicit formula for discriminator in this example reads:
\begin{equation}
\ket{\psi_0} = \frac{1}{\sqrt{2}} (\ket{00} + \ket{11}),
\end{equation}
with final measurements being equal to
\begin{equation}
V_0 =
\left(\begin{array}{cc} \alpha & -\beta\\ \beta & \alpha \end{array}\right),
\end{equation}
and \begin{equation}
V_1 =
\left(\begin{array}{cc} -\beta & \alpha \\ \alpha & \beta \end{array}\right),
\end{equation}
where \begin{equation}
\alpha = \frac{\sqrt{2 - \sqrt{2}}}{2} = \cos\left( \frac{3}{8} \pi \right),
\end{equation}
\begin{equation}
\beta  = \frac{\sqrt{2  + \sqrt{2}}}{2} = \sin\left( \frac{3}{8} \pi \right).
\end{equation}
To use the above benchmarking scheme in PyQBench, we first need to decompose the unitary matrices
above into sequences of gates.
In standard approach, a circuit \todo[inline]{napisac ze circuit zaczyna w stanie 00, lukasz chcial cos o bramkach dosteonych na urzadzeniach, ktore sa typowe itd}
The circuit taking $\ket{00}$ to the Bell state $\ket{\psi_0}$  comprises the Hadamard gate followed by CNOT gate on both qubits (see Fig.~\ref{fig:discriminator}).
\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/discriminator}
	\caption{Decomposition of the Bell state $\ket{\psi_0}$. }
	\label{fig:discriminator}
\end{figure}
For $V_0$ and $V_1$ observe that $V_0 = \operatorname{RY}\left(\frac{3}{4} \pi \right) $, where $\operatorname{RY}$ is rotation gate around the $Y$ axis defined by
\begin{equation}
\operatorname{RY}(\theta) =
\left(\begin{array}{cc} \cos\frac{\theta}{2} & -\sin\frac{\theta}{2} \\ \sin\frac{\theta}{2} & \cos\frac{\theta}{2} \end{array}\right).
\end{equation}
To obtain $V_1$  we need only to swap the columns, i.e.
\begin{equation}
V_1 = \operatorname{RY}\left(\frac{3}{4} \pi \right) \operatorname{X}.
\end{equation}
Finally, the optimal probability of correct discrimination is equal to
\begin{equation}
p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{1}{2} + \frac{\sqrt{2}}{4}.
\end{equation}
We will now demonstrate how to implement this theoretical scheme in PyQBench. For this example we
will use the Qiskit Aer simulator \cite{aer}. First, we import the necessary functions and classes from
PyQBench and Qiskit. We also import \texttt{numpy} for the definition of \texttt{np.pi} constant and
the exponential function. The exact purpose of the imported functions will be described at the point
of their usage.
\begin{lstlisting}[language=Python, caption=Imports needed for running benchmarking example]
import numpy as np
from qiskit import QuantumCircuit, Aer
from qbench.schemes.postselection import benchmark_using_postselection
from qbench.schemes.direct_sum import benchmark_using_direct_sum
\end{lstlisting}
To implement the discrimination scheme in PyQBench, we need to define all the necessary components
as Qiskit instructions. We can do so by constructing a circuit object acting on qubits 0 and 1 and
then converting them using \texttt{to\_instruction()} method.
\begin{lstlisting}[language=Python, caption= Defining components for Hadamard experiment]
def state_prep():
	circuit = QuantumCircuit(2)
	circuit.h(0)
	circuit.cnot(0, 1)
	return circuit.to_instruction()

>>>>

\subsubsection{Importance of choosing the optimal discrimination scheme}
In principle, the schemes described in the previous section could be used with any choice of
$\ket{\psi_0}$ and final measurements $\PP_{V_i}$. However, we argue that it is best to choose those
components in such a way that they maximize the probability of correct discrimination. To see that,
suppose that some choice of $\ket{\psi_0}, \PP_{V_0}, \PP_{V_1}$ yields the theoretical upper bound
of discriminating between two measurements of one, i.e. on a perfect quantum computer you will always
make a correct guess. Then, on real hardware, we might obtain any empirical value in range $\left[\frac{1}{2},
1\right]$. On the other hand, if we choose the components of our scheme such that the successful
discrimination probability is only $\frac{3}{5}$, the possible range of empirically obtainable probabilities
is only $\left[\frac{1}{2}, \frac{3}{5}\right]$. Hence, in the second case, the discrepancy between theoretical and empirical
results will be less pronounced.

\subsubsection{Constructing optimal discrimination scheme}
To construct the optimal discrimination scheme, one starts by calculating the probability of correct discrimination. Using the celebrated result by Helstrom~\cite{helstrom1976quantum}, one finds that the  optimal probability of correct discrimination between two quantum measurements, $\PP$  and $\mathcal{Q}$, is
\begin{equation}
p_{\text{succ}}(\PP, \mathcal{Q}) =  \frac12 + \frac14 \| \PP - \mathcal{Q} \|_\diamond,
\end{equation}
where
\begin{equation}
\|\PP - \QQ\|_\diamond = \max_{\| \ket{\psi}\|_1=1} \| \left( (\PP - \QQ) \otimes \1\right) (\proj{\psi}) \|_1.
\end{equation}
The quantum state $\ket{\psi_0}$ maximizing the diamond norm above is called the
\emph{discriminator}, and can be computed e.g. using semidefinite programming (SDP)
\cite{watrous,watrous2021simplier}. Furthermore,  using the proof of the Holevo-Helstrom theorem, it is
possible to construct corresponding unitaries $V_0$, $V_1$ to create the optimal
discrimination strategy. For brevity, we do not describe this procedure here. Instead, we refer the
interested reader to \cite{watrous}.

@@

\subsection{Using user-defined measurement with \texttt{qbench} package}
In this example, we will demonstrate how \texttt{qbench} package can be used with user--defined
measurement. For this purpose, we will use $U = H$ (the Hadamard gate). The detailed calculations
that lead to the particular form of the discriminator and final measurements can be found in
\ref{app:hadamard}.
\noindent
The explicit formula for discriminator in this example reads:
\begin{equation}
\ket{\psi_0} = \frac{1}{\sqrt{2}} (\ket{00} + \ket{11}),
\end{equation}
with final measurements being equal to
\begin{equation}
V_0 =
\left(\begin{array}{cc} \alpha & -\beta\\ \beta & \alpha \end{array}\right),
\end{equation}
and \begin{equation}
V_1 =
\left(\begin{array}{cc} -\beta & \alpha \\ \alpha & \beta \end{array}\right),
\end{equation}
where \begin{equation}
\alpha = \frac{\sqrt{2 - \sqrt{2}}}{2} = \cos\left( \frac{3}{8} \pi \right),
\end{equation}
\begin{equation}
\beta  = \frac{\sqrt{2  + \sqrt{2}}}{2} = \sin\left( \frac{3}{8} \pi \right).
\end{equation}
To use the above benchmarking scheme in PyQBench, we first need to construct circuits that can be executed by actual hardware. To this end, we need to represent each of the unitaries as a sequence of standard gates, keeping in mind that quantum circuits start execution from  $\ket{00}$ state. The circuit taking $\ket{00}$ to the Bell state $\ket{\psi_0}$  comprises the Hadamard gate followed by CNOT gate on both qubits (see Fig.~\ref{fig:discriminator}).
\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/discriminator}
	\caption{Decomposition of the Bell state $\ket{\psi_0}$. }
	\label{fig:discriminator}
\end{figure}
For $V_0$ and $V_1$ observe that $V_0 = \operatorname{RY}\left(\frac{3}{4} \pi \right) $, where $\operatorname{RY}$ is rotation gate around the $Y$ axis defined by
\begin{equation}
\operatorname{RY}(\theta) =
\left(\begin{array}{cc} \cos\frac{\theta}{2} & -\sin\frac{\theta}{2} \\ \sin\frac{\theta}{2} & \cos\frac{\theta}{2} \end{array}\right).
\end{equation}
To obtain $V_1$  we need only to swap the columns, i.e.
\begin{equation}
V_1 = \operatorname{RY}\left(\frac{3}{4} \pi \right) \operatorname{X}.
\end{equation}
Finally, the optimal probability of correct discrimination is equal to
\begin{equation}
p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{1}{2} + \frac{\sqrt{2}}{4}.
\end{equation}
We will now demonstrate how to implement this theoretical scheme in PyQBench. For this example we
will use the Qiskit Aer simulator \cite{aer}. First, we import the necessary functions and classes from
PyQBench and Qiskit. We also import \texttt{numpy} for the definition of \texttt{np.pi} constant and
the exponential function. The exact purpose of the imported functions will be described at the point
of their usage.
\begin{lstlisting}[language=Python, caption=Imports needed for running benchmarking example]
import numpy as np
from qiskit import QuantumCircuit, Aer
from qbench.schemes.postselection import benchmark_using_postselection
from qbench.schemes.direct_sum import benchmark_using_direct_sum
\end{lstlisting}
To implement the discrimination scheme in PyQBench, we need to define all the necessary components
as Qiskit instructions. We can do so by constructing a circuit object acting on qubits 0 and 1 and
then converting them using \texttt{to\_instruction()} method.
\begin{lstlisting}[language=Python, caption= Defining components for Hadamard experiment]
def state_prep():
	circuit = QuantumCircuit(2)
	circuit.h(0)
	circuit.cnot(0, 1)
	return circuit.to_instruction()
